\chapter*{Abstract}
The goal of this master thesis is to present a parallel approach of the Proper Orthogonal Decomposition (POD) and to integrate it into a reduced order modeling framework.

In the reduced-order modeling community, the POD is a popular snapshot-based method to obtain reduced basis functions.
%Reduced basis methods are projection-based model order reduction techniques to reduce the complexity of parametric differential equations.
In most reduced-order molding frameworks, the computation of the POD is done in serial even though the snapshots are computed on distributed and parallel systems like HPC supercomputers.
This leads to being the bottle-neck in many reduced order modeling frameworks, that can be solved by parallel implementation.

%Parallel
For this master thesis, we took a look at two different approaches to compute the POD in parallel. 
We compared a method using the singular value decomposition to a method solving the eigenvalue problem of the correlation matrix.
The latter method has shown to be the best method in terms of computational cost and speedup.
We were also able to increase the performance of the method by introducing a hybrid implementation combining a multithread with a multi-process paradigm.
The proposed parallel implementation enables the exploitation of HPC supercomputers also for POD computation, and not only for the truth solutions.


%EZyRB
The parallel algorithm is integrated into the reduced order modeling framework EZyRB, a framework for data-driven reduced order modeling, which has been exploited for the numerical results proposed in this work.



 